---
title: "Cape Peninsula Working for Water plots"
author: "Jasper Slingsby"
date: "3 August 2016"
output: html_document
---

```{r, echo=FALSE}

### Get libraries and setwd
suppressPackageStartupMessages(library(vegan))
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressPackageStartupMessages(library(MCMCglmm))

#if (Sys.getenv("USER")=="jasper"){setwd("/Users/jasper/Dropbox/Shared/CapeCommunities")}

dt <- "12July16"

### Get data

##Treatment data
trt <- read.table("Data/LaurePrep/plot_treatments_21June16.txt", header=T)

##Sub-plot data (1x1m)
sp02 <- read.table(paste("Data/LaurePrep/subPlotSpc_2002_NbIndiv_", dt, ".txt", sep=""), stringsAsFactors = F)
sp08 <- read.table(paste("Data/LaurePrep/subPlotSpc_2008_NbIndiv_", dt, ".txt", sep=""), stringsAsFactors = F)
sp11 <- read.table(paste("Data/LaurePrep/subPlotSpc_2011_NbIndiv_", dt, ".txt", sep=""), stringsAsFactors = F)
sp14 <- read.table(paste("Data/LaurePrep/subPlotSpc_2014_NbIndiv_", dt, ".txt", sep=""), stringsAsFactors = F)

#Fix error?
sp02$Wahlenbergia_tenella[which(sp02$Wahlenbergia_tenella==.5)] <- 5

#Fix rownames
r2 <- rownames(sp02)
r8 <- rownames(sp08)
r11 <- rownames(sp11)
r14 <- rownames(sp14)

r2 <- gsub("p", "", r2)
rownames(sp02) <- sapply(r2, function(x) {paste(strsplit(x, split = "_")[[1]][1], which(letters==strsplit(x, split = "_")[[1]][2]), sep="_")})
rownames(sp08) <- sapply(r8, function(x) {paste(strsplit(x, split = "\\.")[[1]][1], strsplit(x, split = "\\.")[[1]][3], sep="_")})
rownames(sp11) <- sapply(r11, function(x) {paste(strsplit(x, split = "\\.")[[1]][1], strsplit(x, split = "\\.")[[1]][3], sep="_")})
rownames(sp14) <- sapply(r14, function(x) {paste(strsplit(x, split = "\\.")[[1]][1], which(LETTERS==strsplit(x, split = "\\.")[[1]][3]), sep="_")})

alien <- read.table("Data/LaurePrep/Exotic_species_list_CapeCom_13July16.txt", header=T)

asp02 <- sp02[,which(colnames(sp02)%in%alien[,1])]
asp08 <- sp08[,which(colnames(sp08)%in%alien[,1])]
asp11 <- sp11[,which(colnames(sp11)%in%alien[,1])]
asp14 <- sp14[,which(colnames(sp14)%in%alien[,1])]

sp02 <- sp02[,-which(colnames(sp02)%in%alien[,1])]
sp08 <- sp08[,-which(colnames(sp08)%in%alien[,1])]
sp11 <- sp11[,-which(colnames(sp11)%in%alien[,1])]
sp14 <- sp14[,-which(colnames(sp14)%in%alien[,1])]

## Create dataframe of rarified "richness", etc
spd2<-rowSums(decostand(sp02,"pa"))
spd8<-rowSums(decostand(sp08,"pa"))
spd11<-rowSums(decostand(sp11,"pa"))
spd14<-rowSums(decostand(sp14,"pa"))

sp02_5 <- sp02[which(spd2>5),]
sp02_5 <- sp02_5[,which(colSums(sp02_5)>0)]
sp08_5 <- sp08[which(spd8>5),]
sp08_5 <- sp08_5[,which(colSums(sp08_5)>0)]
sp11_5 <- sp11[which(spd11>5),]
sp11_5 <- sp11_5[,which(colSums(sp11_5)>0)]
sp14_5 <- sp14[which(spd14>5),]
sp14_5 <- sp14_5[,which(colSums(sp14_5)>0)]

spr2_5 <- rarefy(sp02_5,4)
spr8_5 <- rarefy(sp08_5,4)
spr11_5 <- rarefy(sp11_5,4)
spr14_5 <- rarefy(sp14_5,4)

spd2_5 <- rowSums(decostand(sp02_5,"pa"))
spd8_5 <- rowSums(decostand(sp08_5,"pa"))
spd11_5 <- rowSums(decostand(sp11_5,"pa"))
spd14_5 <- rowSums(decostand(sp14_5,"pa"))

Ind2_5 <- rowSums(sp02_5)
Ind8_5 <- rowSums(sp08_5)
Ind11_5 <- rowSums(sp11_5)
Ind14_5 <- rowSums(sp14_5)

dat <- data.frame(
  Subplot = c(rownames(sp02_5), rownames(sp08_5), rownames(sp11_5), rownames(sp14_5)),
  Year = c(rep(2002, length(spr2_5)), rep(2008, length(spr8_5)), rep(2011, length(spr11_5)), rep(2014, length(spr14_5))),
  Individuals = c(Ind2_5, Ind8_5, Ind11_5, Ind14_5),
  Species_density = c(spd2_5, spd8_5, spd11_5, spd14_5),
  Species_richness = c(spr2_5, spr8_5, spr11_5, spr14_5)
  )

dat$Site <- as.factor(sapply(as.character(dat$Subplot), function(x) {strsplit(x, split="_")[[1]][1]}))

dat <- merge(dat, trt)

dat$Year <- as.factor(dat$Year)
dat$Aliens <- relevel(dat$Aliens, "No_Aliens")

###

##Get unique names
#nms <- sort(unique(c(colnames(sp02), colnames(sp08), colnames(sp11), colnames(sp14))))
#nms <- gsub("_", " ", nms)

#Get trait data
#tr <- read.csv("/Users/jasper/GIT/Dimensions-ZA/data_base/Spec_Trait_All.csv", stringsAsFactors=F)
#trnms <- sort(unique(tr$SpeciesID))

#Match trait and survey data
#length(which(nms %in% trnms))
#nms <- nms[which(nms %in% trnms)]
#nms <- gsub(" ", "_", nms)
#write.table(nms, file="/Users/jasper/Dropbox/Shared/CapeCommunities/Data/spp_matched_to_quantitative_trait_data.txt", sep="\t", quote=F)

```

Let's have a look at the structure of the dataset:

```{r, echo=FALSE}
summary(dat)

hist(table(dat$Subplot))
```

Let's have a look at the relationship between number of individuals, species density and species richness (NEED TO EXCLUDE ALIEN SPECIES FIRST):

```{r, echo=FALSE}
pairs(dat[,c("Individuals", "Species_density", "Species_richness")])
```

And differences between treatments by year?

<br>

All plots:

<br>

```{r, echo=FALSE}
boxplot(Species_density ~ Year + Aliens, data = dat, cex.axis=.75, las=2, ylab = "Species Density")
boxplot(Species_richness ~ Year + Aliens, data = dat, cex.axis=.75, las=2, ylab = "Species Richness")
boxplot(Individuals ~ Year + Aliens, data = dat, cex.axis=.75, las=2, ylab = "Number of individuals")
```

<br>

TMS Fynbos plots only:

<br>

```{r, echo=FALSE}
boxplot(Species_density ~ Year + Aliens, data = dat[which(dat$Geology=="TMS" & dat$Vegetation=="Fynbos"),], cex.axis=.75, las=2, ylab = "Species Density")
boxplot(Species_richness ~ Year + Aliens, data = dat[which(dat$Geology=="TMS" & dat$Vegetation=="Fynbos"),], cex.axis=.75, las=2, ylab = "Species Richness")
boxplot(Individuals ~ Year + Aliens, data = dat[which(dat$Geology=="TMS" & dat$Vegetation=="Fynbos"),], cex.axis=.75, las=2, ylab = "Individuals")
```

<br>

Run MCMCglmm model based on a poisson distribution with subplot as a fixed effect on **species density** across all plots:

<br>

```{r, echo=F}
prior<-list(R=list(V=1, nu=0.002), G=list(G1=list(V=1, nu=0.002)))

MCmod1=MCMCglmm(Species_density ~ Aliens + Year, random=~Subplot, data=dat, family="poisson", prior=prior, verbose=FALSE, pr=TRUE)
summary(MCmod1)
#plot(MCmod1)
```

<br>

Run MCMCglmm model based on a gaussian distribution with subplot as a fixed effect on **species richness** across all plots:

<br>

```{r, echo=F}
prior<-list(R=list(V=1, nu=0.002), G=list(G1=list(V=1, nu=0.002)))

MCmod1=MCMCglmm(Species_richness ~ Aliens + Year, random=~Subplot, data=dat, family="gaussian", prior=prior, verbose=FALSE, pr=TRUE)
summary(MCmod1)
#plot(MCmod1)
```

<br>

Notice that both analyses show impacts in invaded and cleared sites, but the key difference is that species density is lower in each year relative to 2002, while species richness is actually higher with each year relative to 2002!

<br>

Now lets look at TMS Fynbos plots only:

<br>

**Species density:**

<br>

```{r, echo=F}
prior<-list(R=list(V=1, nu=0.002), G=list(G1=list(V=1, nu=0.002)))

MCmod1=MCMCglmm(Species_density ~ Aliens + Year, random=~Subplot, data= dat[which(dat$Geology=="TMS" & dat$Vegetation=="Fynbos"),], family="poisson", prior=prior, verbose=FALSE, pr=TRUE)
summary(MCmod1)
#plot(MCmod1)
```

<br>

**Species richness:**

<br>

```{r, echo=F}
prior<-list(R=list(V=1, nu=0.002), G=list(G1=list(V=1, nu=0.002)))

MCmod1=MCMCglmm(Species_richness ~ Aliens + Year, random=~Subplot, data= dat[which(dat$Geology=="TMS" & dat$Vegetation=="Fynbos"),], family="gaussian", prior=prior, verbose=FALSE, pr=TRUE)
summary(MCmod1)
#plot(MCmod1)
```

<br>

Same story...

<br>


